Simulating High-Dimensional Multivariate Data

### 2021-02-24 040-monte-carlo-evaluation.Rmd
updated all bigsimr code
moved generated data into separate scripts

```{r ch040-LoadLib040, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, results = "hide"}
## devtools::install_github("SchisslerGroup/bigsimr", ref="dev-julia")
library(bigsimr)
Sys.setenv(JULIA_NUM_THREADS=parallel::detectCores())
## Either let JuliaCall find your installation of Julia or set manually
# bs <- bigsimr_setup(JULIA_HOME = "/usr/local/bin/")
bs <- bigsimr_setup(JULIA_HOME = "/home/alex/julia-1.5.3/bin/", pkg_check = FALSE)
dist <- distributions_setup() # Julia distributions
library(tidyverse)
cores <- as.integer( parallel::detectCores() - 1 )
## cores <- as.integer( 1 )
set.seed(10162020)
```

@@

```{r ch040-preview, echo = FALSE, eval = FALSE}
bookdown::preview_chapter('040-monte-carlo-evaluation.Rmd')
```

@@

```{r ch040-biNormal, echo = FALSE, eval = TRUE}
mu <- 0
sigma <- 1
if (!file.exists( 'results/biNorm_sims.rds'))  {
    mom_norm <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(mean = m, sd = s)
    }
    mu <- 0
    sigma <- 1
    margins <- c(
        dist$Normal(mu, sigma),
        dist$Normal(mu, sigma)
    )
    types <- c("Pearson", "Spearman", "Kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)
    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n, stringsAsFactors = FALSE)
    res <- data.frame()
    dir.create("./results/biNorm_sims", showWarnings = TRUE, recursive = TRUE)
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]
        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)
        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)
            time_data <- system.time({
                ## x <- rmvn(n = n, mu = rep( mu, 2 ), sigma = Rho)
                ## use rvec to test our software
                x <- bs$rvec(n, Rho, margins)
            })
            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "1CORE",
                "-lib", "bigsimr"
            )
            ## save sims in their own subdirectory
            saveRDS(x, file = paste0("./results/biNorm_sims/", id, ".rds"))
            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            norm_args_hat <- mom_norm(x[,1])
            mu_hat <- norm_args_hat$mean
            sigma_hat <- norm_args_hat$sd
            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "norm",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  mean = mu,
                                  sd = sigma,
                                  mean_hat = mu_hat,
                                  sd_hat = sigma_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNorm_sims.rds")
}
```

```{r ch040-biGamma, echo = FALSE, eval = TRUE}
shape <- 10
rate <- 1
if (!file.exists( 'results/biGamma_sims.rds'))  {
    mom_gamma <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(shape = m^2 / s^2, rate = m / s^2)
    }
    margins <- c(
        dist$Gamma(shape, rate),
        dist$Gamma(shape, rate)
    )

    types <- c("Pearson", "Spearman", "Kendall")
    ## cores <- c(1L)
    cores <- as.integer( parallel::detectCores() - 1 )
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n,
                            stringsAsFactors = FALSE)

    dir.create("./results/biGamma_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {

        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]

        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        ## rho = cor_seq[1]
        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)

            time_data <- system.time({
                x <- bs$rvec(n, Rho, margins)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biGamma_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            gamma_args_hat <- mom_gamma(x[,1])
            shape_hat <- gamma_args_hat$shape
            rate_hat <- gamma_args_hat$rate

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "gamma",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  shape = shape,
                                  rate = rate,
                                  shape_hat = shape_hat,
                                  rate_hat = rate_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biGamma_sims.rds")
}
```

```{r ch040-biNegBin, echo = FALSE, eval = TRUE}
size <- 4
prob <- 3e-04
if (!file.exists( 'results/biNB_sims.rds'))  {
    mom_nbinom <- function(x) {
        m <- mean(x)
        s <- sd(x)
        list(size = m^2 / (s^2 - m), prob = m / s^2)
    }

    margins <- c(
        dist$NegativeBinomial(size, prob),
        dist$NegativeBinomial(size, prob)
    )

    types <- c("Pearson", "Spearman", "Kendall")
    cores <- as.integer( parallel::detectCores() - 1 )
    ## cores <- cores
    n <- c(1e3, 1e4, 1e5)
    ## n <- c(1e1, 1e2, 1e3)

    eps <- 1e-2
    grid_steps <- 100
    ## grid_steps <- 10
    sim_pars <- expand.grid(type = types, cores = cores, n = n,
                            stringsAsFactors = FALSE)

    dir.create("./results/biNB_sims", showWarnings = TRUE, recursive = TRUE)
    res <- data.frame()
    for (i in 1:nrow(sim_pars)) {
        type <- sim_pars$type[i]
        cores <- sim_pars$cores[i]
        n <- sim_pars$n[i]

        tmp_bounds <- bs$cor_bounds(margins, bs[[type]])
        cor_lo <- tmp_bounds$lower[1,2] + eps
        cor_hi <- tmp_bounds$upper[1,2] - eps
        cor_seq <- seq(cor_lo, cor_hi, length.out = grid_steps)

        for (rho in cor_seq) {
            Rho <- matrix(rho, 2, 2)
            diag(Rho) <- 1.0
            Rho <- bs$cor_convert(Rho, bs[[type]], bs$Pearson)

            time_data <- system.time({
                x <- bs$rvec(n, Rho, margins)
            })

            ## Save the sims in case
            id <- paste0(
                "d", 2,
                "-N", n,
                "-c", cores,
                "-r", rho,
                "-Cor", type,
                "-dev", "CPU",
                "-lib", "bigsimr"
            )
            saveRDS(x, file = paste0("./results/biNB_sims/", id, ".rds"))

            ## Estimate statistics
            Rho_hat <- bs$cor_fast(x, bs[[type]])
            rho_hat <- Rho_hat[1, 2]
            nbinom_args_hat <- mom_nbinom(x[,1])
            size_hat <- nbinom_args_hat$size
            prob_hat <- nbinom_args_hat$prob

            ## Save the results
            res <- rbind(res, data.frame(
                                  method = "bigsimr",
                                  device = "CPU",
                                  type = type,
                                  cores = cores,
                                  margins = "nbinom",
                                  d = 2,
                                  N = n,
                                  rho = rho,
                                  rho_hat = rho_hat,
                                  size = size,
                                  prob = prob,
                                  size_hat = size_hat,
                                  prob_hat = prob_hat,
                                  sim_time = unname(time_data["elapsed"])
                              ))
        }
    }
    res$type <- factor(res$type, levels = c("pearson",  "spearman", "kendall" ) )
    saveRDS(object = res, "./results/biNB_sims.rds")
}
```

Table \@ref(tab:sims) lists our identical-marginal, bivariate simulation configurations.
We increase the simulate replicates $B$ to check that our results converge to the target correlations and gauge statistical efficiency.
We select distributions beginning with a standard multivariate normal (MVN) as we expect the performance to be exact (up to MC error) for all correlation types.
Then, we select a non-symmetric continuous distribution: a standard (rate =1) two-component multivariate gamma (MVG).
Finally, we select distributions and marginal parameter values that are motivated by our RNA-seq data, namely values proximal to probabilities and sizes estimated from the data (see [Example applications](examples) for estimation details).
Thus we arrive at a multivariate negative binomial (MVNB) $p_1 = p_2 = `r prob`, r_1 = r_2 = 4,\rho \in \Omega$.

@@

*Bivariate Gamma*. 
Similarly, let's check the performance for a non-symmetric continuous distribution: a standard (rate =1) bivariate gamma.
Here we have a Bivariate Gamma with $shape_1 = shape_2 = 10, \rho_{type}$.
We vary $\rho$ across the entire possible range of correlations for each correlation type.
Figure \@ref(fig:ch040-biGammaPlot) `bigsimr` recovers the Pearson specified correlations for bivariate gamma.
*Bivariate Negative Binomial*. 
Figure \@ref(fig:ch040-biNegBinPlot) `bigsimr` recovers the Pearson specified correlations for bivariate negative binomial.

 -->

```{r ch040-biNormPlot, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, fig.height = 5, fig.width= 8, fig.align='center', fig.cap = "`bigsimr` recovers the Pearson specified correlations for MVN."}
dat <- readRDS("./results/biNorm_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biNormPlot.pdf')
```

```{r ch040-biGammaPlot, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, fig.height = 5, fig.width= 8, fig.align='center', fig.cap = "`bigsimr` recovers the Pearson specified correlations for Bivariate Gamma."}
dat <- readRDS("./results/biGamma_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biGammaPlot.pdf')
```

```{r ch040-biNegBinPlot, echo = FALSE, eval = FALSE, message=FALSE, warning=FALSE, fig.path='fig/plot-', dev='png', fig.ext='png', fig.align='center', fig.width = 8, fig.height=5,fig.cap = "`bigsimr` recovers the correlations for bivariate negative binomial only approximately for Pearson but (nearly) exactly for the rank-based correlations."}
dat <- readRDS("./results/biNB_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biNegBinPlot.pdf')
```

```{r ch040-combineBiSims, cache=F}
## compare relative differences in MSE for each dependency, compared to Pearson
allDat <- NULL
dat <- readRDS("./results/biNorm_sims.rds") %>% select( margins, type, N, rho, rho_hat  ) 
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biGamma_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biNB_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
allDat$margins <- factor(allDat$margins, levels = c( 'norm', 'gamma', 'nbinom'),
                         labels = c( 'MVN', 'MVG', 'MVNB' ) )
allDat$type <- factor(allDat$type, levels = c( 'pearson', 'spearman', 'kendall' ),
                      labels = c( 'Pearson', 'Spearman', 'Kendall' ) )
allDat$N <- factor(allDat$N, levels = c( 1000, 10000, 100000) )
```

@@

```{r ch040-BiError, echo = FALSE, eval=TRUE, message=FALSE, warning=FALSE}
## tabMSE  <- allDat %>%
##    group_by( N, margins ) %>%
##    summarize( mse = mean( ( rho - rho_hat )^2 ) )
tabMAE  <- allDat %>%
    ## filter( N == 100000) %>%
    group_by( N, type, margins ) %>%
##     summarize( mse = mean( ( rho - rho_hat )^2 ) ) %>%
    summarize( MAE = mean( abs( rho - rho_hat )  ) )
## normMSE <- as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ] ) 
## ( tabMSE$mse - normMSE ) / normMSE
## ( tabMSE$mse - as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ]) )
## ( tabMSE$mse - as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ]) ) / tabMSE$mse

@@

```{r ch040-bPlot, cache=F, echo = FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.asp = 1.30, fig.weight = 4, fig.align='center', fig.cap = "Bivariate simulations match specified correlations. The horizontal axis plots the specified target correlations across the entire range of possible correlations for each bivariate margin. Normal margins are plotted in dark blue, gamma in medium blue, and negative binomial in light blue. As the number of simulated vectors $B$ increases from left to right, the variation in estimated correlations (vertical axis) decreases. The dashed line indicates equality between the specified and estimated correlations. Only the normal margins match the Pearson correlations (top row) exactly across the entire range of correlations. The two non-normal bivariate random vectors experience attenuation (downward bias) for extreme Pearson correlations (due to limitations in our algorithm) and more restriction (due to the Frechet bounds). The rank-based correlations (bottom two rows) are matched exactly for all margins and obtain the full range of possible dependencies."}
# https://www.datanovia.com/en/blog/how-to-change-ggplot-facet-labels/
# New facet label names
repsLabs <- paste0("B=", c("1,000", "10,000", "100,000") )
names(repsLabs) <- c(1000, 10000, 100000)
typeLabs <- c( 'Pearson', 'Spearman', 'Kendall' )
names(typeLabs) <- c( 'Pearson', 'Spearman', 'Kendall' )

@@

<!-- 
See [Discussion](discussion) for future directions for fast Pearson matching and discrete-specific modifications.
your comment -->

>>>>

@@

```{r ch040-LoadLib040, include=FALSE}
library(tidyverse)
library(bigsimr)

Sys.setenv(JULIA_NUM_THREADS = parallel::detectCores())
bs <- bigsimr_setup()
dist <- distributions_setup()

cores <- as.integer(parallel::detectCores() - 1)
set.seed(10162020)
```

@@

Table \@ref(tab:sims) lists our identical-marginal, bivariate simulation configurations. We increase the simulate replicates $B$ to check that our results converge to the target correlations and gauge statistical efficiency. We select distributions beginning with a standard multivariate normal (MVN) as we expect the performance to be exact (up to MC error) for all correlation types. Then, we select a non-symmetric continuous distribution: a standard (rate =1) two-component multivariate gamma (MVG). Finally, we select distributions and marginal parameter values that are motivated by our RNA-seq data, namely values proximal to probabilities and sizes estimated from the data (see [Example applications](examples) for estimation details). Thus we arrive at a multivariate negative binomial (MVNB) $p_1 = p_2 = 3e-04, r_1 = r_2 = 4, \rho \in \Omega$.

@@

*Bivariate Gamma*. 
Similarly, let's check the performance for a non-symmetric continuous distribution: a standard (rate =1) bivariate gamma.
Here we have a Bivariate Gamma with $shape_1 = shape_2 = 10, \rho_{type}$.
We vary $\rho$ across the entire possible range of correlations for each correlation type.
Figure \@ref(fig:ch040-biGammaPlot) `bigsimr` recovers the Pearson specified correlations for bivariate gamma.
*Bivariate Negative Binomial*. 
Figure \@ref(fig:ch040-biNegBinPlot) `bigsimr` recovers the Pearson specified correlations for bivariate negative binomial.
 -->

```{r ch040-biNormPlot, eval=FALSE, fig.height=5, fig.width=8, fig.cap="`bigsimr` recovers the Pearson specified correlations for MVN."}
dat <- readRDS("./results/biNorm_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biNormPlot.pdf')
```

```{r ch040-biGammaPlot, eval=FALSE, fig.height=5, fig.width=8, fig.cap="`bigsimr` recovers the Pearson specified correlations for Bivariate Gamma."}
dat <- readRDS("./results/biGamma_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biGammaPlot.pdf')
```

```{r ch040-biNegBinPlot, eval=FALSE, fig.path='fig/plot-', dev='png', fig.ext='png', fig.width=8, fig.height=5, fig.cap="`bigsimr` recovers the correlations for bivariate negative binomial only approximately for Pearson but (nearly) exactly for the rank-based correlations."}
dat <- readRDS("./results/biNB_sims.rds")
dat %>%
    filter(cores == cores) %>%
    ggplot(aes(rho, rho_hat, color = type)) +
    ## ggplot(aes(rho, rho_hat)) +
    geom_point() +
    geom_abline(slope = 1) +
    ## facet_wrap(~ + N) + theme_bw()
    facet_wrap(~ type + N) + theme_bw()
## ggsave('fig/plot-biNegBinPlot.pdf')
```

```{r ch040-combineBiSims, cache=FALSE}
## compare relative differences in MSE for each dependency, compared to Pearson
allDat <- NULL
dat <- readRDS("./results/biNorm_sims.rds") %>% select( margins, type, N, rho, rho_hat  ) 
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biGamma_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
dat <- readRDS("./results/biNB_sims.rds") %>% select( margins, type, N, rho, rho_hat  )
allDat <- rbind( allDat, dat )
allDat$margins <- factor(allDat$margins, levels = c( 'norm', 'gamma', 'nbinom'),
                         labels = c( 'MVN', 'MVG', 'MVNB' ) )
allDat$type <- factor(allDat$type, levels = c( 'pearson', 'spearman', 'kendall' ),
                      labels = c( 'Pearson', 'Spearman', 'Kendall' ) )
allDat$N <- factor(allDat$N, levels = c( 1000, 10000, 100000) )
```

@@

```{r ch040-BiError}
## tabMSE  <- allDat %>%
##    group_by( N, margins ) %>%
##    summarize( mse = mean( ( rho - rho_hat )^2 ) )
tabMAE  <- allDat %>%
    ## filter( N == 100000) %>%
    group_by( N, type, margins ) %>%
##     summarize( mse = mean( ( rho - rho_hat )^2 ) ) %>%
    summarize( MAE = mean( abs( rho - rho_hat )  ) )
## normMSE <- as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ] ) 
## ( tabMSE$mse - normMSE ) / normMSE
## ( tabMSE$mse - as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ]) )
## ( tabMSE$mse - as.numeric( tabMSE[ tabMSE$margins == 'norm', 'mse' ]) ) / tabMSE$mse

@@

```{r ch040-bPlot, cache=FALSE, fig.asp=1.30, fig.weight=4, fig.cap = "Bivariate simulations match specified correlations. The horizontal axis plots the specified target correlations across the entire range of possible correlations for each bivariate margin. Normal margins are plotted in dark blue, gamma in medium blue, and negative binomial in light blue. As the number of simulated vectors $B$ increases from left to right, the variation in estimated correlations (vertical axis) decreases. The dashed line indicates equality between the specified and estimated correlations. Only the normal margins match the Pearson correlations (top row) exactly across the entire range of correlations. The two non-normal bivariate random vectors experience attenuation (downward bias) for extreme Pearson correlations (due to limitations in our algorithm) and more restriction (due to the Frechet bounds). The rank-based correlations (bottom two rows) are matched exactly for all margins and obtain the full range of possible dependencies."}
# https://www.datanovia.com/en/blog/how-to-change-ggplot-facet-labels/
# New facet label names
repsLabs <- paste0("B=", c("1,000", "10,000", "100,000") )
names(repsLabs) <- c(1000, 10000, 100000)
typeLabs <- c( 'Pearson', 'Spearman', 'Kendall' )
names(typeLabs) <- c( 'Pearson', 'Spearman', 'Kendall' )

@@

<!-- 
See [Discussion](discussion) for future directions for fast Pearson matching and discrete-specific modifications.
-->

### 2021-02-24 060-conclusion-discussion.Rmd
updated all bigsimr code
moved generated data into separate scripts

Researchers working in multivariate computation frequently encounter such difficulties and need to find a close positive (semi-)definite matrix.
A widely-available routine for this task in `R` is `matrix::nearPD`, though it is not suitable for high dimensions.
To overcome this issue, we introduce `bigsimr::nearPSD` --- an implementation of @QS2006's method that computes the nearest correlation matrix to a given symmetric matrix in the Frobenius norm.
This routine as many applications beyond our primary goal of random vector generation.

@@

<!-- 
Future work includes developing scaleable algorithms to match the Pearson correlation matrix more precisely, discrete-margin specific modifications including fast Spearman's correlation rescaling (see Equation \@ref(eq:spearmanRescaled)), and high-dimensional covariance estimation.
From an implementation standpoint, `bigsimr` only supports Nvidia GPUs and redesigning the code using OpenCL would broaden the users who would benefit.
As data-analytic problems grow to even larger dimension, multi-GPU support is a promising hardware-based future direction.
your comment -->

>>>>

Researchers working in multivariate computation frequently encounter such difficulties and need to find a close positive (semi-)definite matrix. A widely-available routine for this task in `R` is `matrix::nearPD`, though it is not suitable for high dimensions. To overcome this issue, we introduce `cor_nearPD` --- an implementation of @QS2006's method that computes the nearest correlation matrix to a given symmetric matrix in the Frobenius norm. This routine as many applications beyond our primary goal of random vector generation.

@@

<!-- 
Future work includes developing scaleable algorithms to match the Pearson correlation matrix more precisely, discrete-margin specific modifications including fast Spearman's correlation rescaling (see Equation \@ref(eq:spearmanRescaled)), and high-dimensional covariance estimation. From an implementation standpoint, `bigsimr` only supports Nvidia GPUs and redesigning the code using OpenCL would broaden the users who would benefit. As data-analytic problems grow to even larger dimension, multi-GPU support is a promising hardware-based future direction.
-->

### 2021-02-24 070-supplementary-materials.Rmd
updated all bigsimr code
moved generated data into separate scripts

```{r, out.width='5%', fig.align='center'}
knitr::include_graphics('images/hex-bigsimr.png')
```

@@

<!--

Nomenclature/Notation

```{r ch070-all-dists, eval=FALSE}
all_dists <- list(
  list(dist = "beta", shape1, shape2),
  list(dist = "binom", size, prob),
  list(dist = "cauchy", location, scale),
  list(dist = "chisq", df),
  list(dist = "exp", rate),
  list(dist = "f", df1, df2),
  list(dist = "gamma", shape, rate),
  list(dist = "geom", prob),
  list(dist = "hyper", m, n, k),
  list(dist = "logis", location, scale),
  list(dist = "lnorm", meanlog, sdlog),
  list(dist = "nbinom", size, prob),
  list(dist = "norm", mean, sd),
  list(dist = "pois", lambda),
  list(dist = "t", df),
  list(dist = "unif", min, max),
  list(dist = "weibull", shape, scale),
  list(dist = "wilcox", m, n),
  list(dist = "signrank", n)
)
```

your comment -->

>>>>

```{r, out.width='5%'}
knitr::include_graphics('images/hex-bigsimr.png')
```

@@

<!-- 
Nomenclature/Notation
-->
